\LARGE \bf Steering a Swarm of Particles\\ Using Global Inputs and Swarm Statistics

### 2017-11-16 2017-tro-SteeringSwarmParticles/SwarmControlJournal.tex
Added all the pictures and bios and shrinked the size.etc

\maketitle
\thispagestyle{empty}
\pagestyle{empty}
\begin{abstract}
Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion.

Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
To identify the key parameters for particle manipulation, we used a collection of online games where players steer swarms of up to 500 particles to complete manipulation challenges. We recorded statistics from over ten thousand players. 
Inspired by techniques where human operators performed well, we investigate controllers that use only the mean and variance of the swarm. We prove the mean position is controllable and provide conditions under which variance is controllable.  We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the particle distribution. 
Finally, we employ these controllers as primitives for an object manipulation task and implement all controllers on 100 kilobots controlled by the direction of a global light source.
\end{abstract}

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{shiva.jpg}}]{Shiva Shahrokhi}(S`14) is a PhD Candidate in the Electrical and Computer Engineering Department at the University of Houston expected to graduate in 2018.

She received the B.S. degree in Computer Engineering from Iran University of Science and Technology,
Tehran, Iran. She is involved in studies of swarm manipulation and control for robotics applications.
These include steering massively under-actuated mobile robotic ensembles, swarm shape control, and machine learning.

\end{IEEEbiography}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Lillian.jpg}}]{Lillian Lin}(S`14) is currently pursuing a B.S. degree in Electrical Engineering from the University of Houston, Houston, Texas, and is expected to graduate in 2018.
She is currently the president of the student IEEE branch on campus. She works as an undergraduate researcher at the University of Houston at the Robotics Swarm Control lab. Her main research interests include robotic control systems.

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{chris_portrait.jpg}}]{Chris Ertel} received his B.S. in mechanical engineering from Rice University in 2010, and has since worked in industry and startups developing games, computer-aided design applications, medical telemetry systems, and other software.
Since 2013 he's collaborated with Dr. Becker on SwarmControl.net and related projects at the University of Houston Robotic Swarm Control Lab and the Rice Multi-Robot Systems Lab.

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mable.jpg}}]{Mable Wan} received a B.S. degree in electrical and computer engineering and a B.B.A. in entrepreneurship from the University of Houston in 2017.
Since May 2017, she has been working with Northrop Grumman Corporation in Baltimore, MD as a software engineer.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{DrB.jpg}}]{Aaron T. Becker}(M`06) received M.S. and PhD degrees in electrical and computer engineering from the University of Illinois at Urbana-Champaign, in 2008 and 2012. He was a post doctoral research scholar at the Multi-Robot Systems Lab at Rice University and a post doctoral research fellow at Boston Children's Hospital, Harvard Medical School, before joining the Electrical and Computer Engineering Department at the University of Houston as an Assistant Professor in 2014. He won the best paper at IROS 2013 and received the NSF CAREER award in 2016. He is active in developing control laws and algorithms for robot swarms.

>>>>

\maketitle
\thispagestyle{empty}
\pagestyle{empty}
\begin{abstract}
Microrobotics has the potential to revolutionize many applications including targeted material delivery, assembly, and surgery.  The same properties that promise breakthrough solutions---small size and large populations---present unique challenges for controlling motion. 
Robotic manipulation usually assumes intelligent agents, not particle systems manipulated by a global signal.
To identify the key parameters for particle manipulation, we used a collection of online games where players steer swarms of up to 500 particles to complete manipulation challenges. We recorded statistics from over ten thousand players. 
Inspired by techniques where human operators performed well, we investigate controllers that use only the mean and variance of the swarm. We prove the mean position is controllable and provide conditions under which variance is controllable.  We next derive automatic controllers for these and a hysteresis-based switching control to regulate the first two moments of the particle distribution. 
Finally, we employ these controllers as primitives for an object manipulation task and implement all controllers on 100 kilobots controlled by the direction of a global light source.
\end{abstract}

\begin{IEEEkeywords}
Multi-robot systems, Robot motion 
\end{IEEEkeywords}

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{shiva.jpg}}]{Shiva Shahrokhi}(S`14) is a PhD Candidate in the Electrical and Computer Engineering Department at the University of Houston expected to graduate in 2018. 
She received the B.S. degree in Computer Engineering from Iran University of Science and Technology,
Tehran, Iran. She is involved in studies of swarm manipulation and control for robotics applications.
These include steering massively under-actuated mobile robotic ensembles, swarm shape control, and machine learning.
\end{IEEEbiography}
\vspace{-2em}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Lillian.jpeg}}]{Lillian Lin}(S`14) is currently pursuing a B.S. degree in Electrical Engineering from the University of Houston, Houston, Texas, and is expected to graduate in 2018.
She is currently the president of the student IEEE branch on campus. She works as an undergraduate researcher at the University of Houston at the Robotics Swarm Control lab. Her main research interests include robotic control systems.
\end{IEEEbiography}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{chris_portrait.jpg}}]{Chris Ertel} received his B.S. in mechanical engineering from Rice University in 2010, and has since worked in industry and startups developing games, computer-aided design applications, medical telemetry systems, and other software.
Since 2013 he's collaborated with Dr. Becker on SwarmControl.net and related projects at the University of Houston Robotic Swarm Control Lab and the Rice Multi-Robot Systems Lab. 
\end{IEEEbiography}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mable.jpg}}]{Mable Wan} received a B.S. degree in electrical and computer engineering and a B.B.A. in entrepreneurship from the University of Houston in 2017.
Since May 2017, she has been working with Northrop Grumman Corporation in Baltimore, MD as a software engineer.
\end{IEEEbiography}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{DrB.jpg}}]{Aaron T. Becker}(S`06-M`12-SM`17) received M.S. and PhD degrees in electrical and computer engineering from the University of Illinois at Urbana-Champaign, in 2008 and 2012. He was a post doctoral research scholar at the Multi-Robot Systems Lab at Rice University and a post doctoral research fellow at Boston Children's Hospital, Harvard Medical School, before joining the Electrical and Computer Engineering Department at the University of Houston as an Assistant Professor in 2014. He won the best paper at IROS 2013 and received the NSF CAREER award in 2016. He is active in developing control laws and algorithms for robot swarms.

### 2017-11-16 2017-tro-SteeringSwarmParticles/experiment.tex
modified the publishing version
Added all the pictures and bios and shrinked the size.etc

The table below summarizes the simulation results for each 10 successful trials:
\begin{center}
\begin{tabular}{ c c }
\hline
Method & Result, mean$\pm$std (s)\\ [0.5ex] 
\hline \hline
Value Iteration (VI) & 367 $\pm$ 253 \\ 
\hline
VI + Potential Fields (PF) & 271 $\pm$ 267 \\  
\hline
VI + Outlier Rejection (OR) & 245 $\pm$ 135\\
\hline
BFS + PF + OR & 183 $\pm$ 179 \\
\hline
VI + PF + OR & 90 $\pm$ 35 \\[0.5ex]
\hline
\end{tabular}
\end{center}
\subsection{Learning a policy for the object}\label{subsec:objectpolicy}

@@

% A corresponding reward function gives a high reward to the goal state,  
% a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 where $x$ is the current state and $\mathbf{u}$ is the action.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is %the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}  \left[ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x, \mathbf{u}) \right].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_i) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_i) = \gamma \max_{\mathbf{u}} \left[r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})\right] \label{eq:ValueIteration} \nonumber\\
\text{end}&  %page 502
\end{align}
Our probabilistic motion model $p(x_j | x, \mathbf{u})$ assumed the object moved in the commanded direction $\mathbf{u}$ half of the time but $+45^\circ$ with probability 0.25 and $-45^\circ$ with probability 0.25. 
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

\begin{figure}
\centering
\renewcommand{\figwid}{3cm}
\begin{overpic}[height=\figwid]{GradientView.png}
\end{overpic}
\begin{overpic}[height=\figwid]{PolicyIter.png}
\end{overpic}
\begin{overpic}[height=\figwid]{MDPmap.pdf}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:BFSGradient} BFS finds the shortest path for the moveable object to compute gradient vectors (left). Modeling as an MDP enables encoding penalties for being near obstacles. (Middle) The control policy from value iteration. (Right) The vision algorithm detects obstacles in the hardware setup. This map is used to produce the value function and control policy shown. 
%\vspace{-2em}
}
\end{figure}

@@

\begin{figure}
\centering
\begin{overpic}[width=1\columnwidth]{PotentialField.jpg}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:potentialField} (Left) The attractive field is centered behind the object's COM. (Middle) The repulsive field is centered at the object's COM. (Right) Combining these forces prevents the swarm from pushing the object backwards.}
\end{figure}

@@

\begin{figure}
\begin{center}
	\begin{overpic}[width=0.9\columnwidth]{StateMachine.pdf}\put(-4,25){\textbf{a}}\end{overpic}\\
	~~\begin{overpic}[width=0.44\columnwidth]{MainRegions.pdf} %,grid=true
	\put(-6,85){\textbf{b}}
	\end{overpic}
	~~
	\begin{overpic}[width=0.48\columnwidth]{TransferRegions.pdf}
	\put(-6,85){\textbf{c}}
	\end{overpic}
\end{center}
\vspace{-0.5em}
\caption{\label{fig:Region}  Outlier rejection state machine and regions.
}
\end{figure}

@@

\begin{figure}
%  \vspace{-20pt}
  \begin{center}
\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\end{overpic}
  \end{center}
%  \vspace{-1em}
\caption{\label{fig:Shapes} The six equal-area objects tested in simulation. %Six object shapes were tested.
\vspace{-1em}
}
\end{figure}

@@

%\subsection{Automated object manipulation (simulation)}
Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation.
 To illustrate the flexibility of the algorithm we tested two additional workspaces, \emph{{\sffamily E}-shaped} and  \emph{Spiral}, without changing the algorithm. These are shown in Fig.~\ref{fig:Emaze}. 
 More complicated workspaces could be generated by composing these workspaces.
 Fig.~\ref{fig:DifferentMazes} shows the results of all  three mazes. The {\sffamily E}-shaped maze required the least average time because the path to the goal is shorter. 
   Experimental results of parameters sweeps are summarized in Fig.~\ref{fig:AutoVeryParam}.  Each trial measured the time to deliver the object to the goal location.  The default parameter settings used 100 particles, a normalized weight of 1, a hexagon shape, and Brownian noise (applied once each simulation step) with $W=5$.

@@

\begin{overpic}[width =\columnwidth]{SwarmRun.pdf}\end{overpic}
\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 particles under automatic control (see also Extension 1).
}
\end{figure}

\begin{figure}\begin{center}\begin{overpic}[width=0.4\columnwidth]{MDPEmaze.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
\vspace{.25em}
\begin{overpic}[width=0.4\columnwidth]{Spiral.pdf}\end{overpic}~\begin{overpic}[width=0.4\columnwidth]{SpiralSuccess.pdf}\end{overpic}
\end{center}
\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
}
\end{figure}
\begin{figure}
\centering
\begin{overpic}[width=.5\columnwidth]{DifferentShapes.pdf}\end{overpic}
\vspace{-1em}
\caption{\label{fig:DifferentMazes} Completion times for three workspaces.}
\end{figure}

@@

\begin{figure*}
\centering
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =\figwid]{SimVeryNum.pdf}\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryNoise.pdf}\put(1,55){b)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryWeight.pdf}\put(1,55){c)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryShape.pdf}\put(1,55){d)}
\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of particles, b) different noise values, c) object weight, and d) object shape.  Each bar is labelled with the number of trials. Completion time is in seconds.
%\vspace{-2em}
}
\end{figure*}

>>>>

The table below summarizes the simulation results for each 10 successful trials.
\begin{center}
\begin{tabular}{ c c }
\hline
Method & Result, mean$\pm$std (s)\\ [0.5ex] 
\hline \hline
Value Iteration (VI) & 367 $\pm$ 253 \\ 
\hline
VI + Potential Fields (PF) & 271 $\pm$ 267 \\  
\hline
VI + Outlier Rejection (OR) & 245 $\pm$ 135\\
\hline
BFS + PF + OR & 183 $\pm$ 179 \\
\hline
VI + PF + OR & 90 $\pm$ 35 \\[0.5ex]
\hline
\end{tabular}
\end{center}
\subsection{Learning a policy for the object}\label{subsec:objectpolicy}

@@

% A corresponding reward function gives a high reward to the goal state,  
% a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 where $x$ is the current state and $\mathbf{u}$ is the action.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is %the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}  \left[ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x, \mathbf{u}) \right].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_i) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_i) = \gamma \max_{\mathbf{u}} \left[r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})\right] \label{eq:ValueIteration} \nonumber\\
\text{end}&  %page 502
\end{align}
Our probabilistic motion model $p(x_j | x, \mathbf{u})$ assumed the object moved in the commanded direction $\mathbf{u}$ half of the time but $+45^\circ$ with probability 0.25 and $-45^\circ$ with probability 0.25. 
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. %A {\sc Matlab} implementation is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

\begin{figure}
\centering
\renewcommand{\figwid}{3cm}
\begin{overpic}[height=\figwid]{GradientView.png}
\end{overpic}
\begin{overpic}[height=\figwid]{PolicyIter.png}
\end{overpic}
\begin{overpic}[height=\figwid]{MDPmap.pdf}
\end{overpic}
\vspace{-0.em}
\caption{\label{fig:BFSGradient} BFS finds the shortest path for the moveable object to compute gradient vectors (left). Modeling as an MDP enables encoding penalties for being near obstacles. (Middle) The control policy from value iteration. (Right) The vision algorithm detects obstacles in the hardware setup. This map is used to produce the value function and control policy shown. 
%\vspace{-2em}
}
\end{figure}

@@

\begin{figure}
\centering
\begin{overpic}[width=1\columnwidth]{PotentialField.jpg}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\vspace{-1.5em}
\caption{\label{fig:potentialField} (Left) The attractive field is centered behind the object's COM. (Middle) The repulsive field is centered at the object's COM. (Right) Combining these forces prevents the swarm from pushing the object backwards.}

\end{figure}

@@

\begin{figure}
\begin{center}
	\begin{overpic}[width=0.9\columnwidth]{StateMachine.pdf}\put(-4,25){\textbf{a}}\end{overpic}\\
	~~\begin{overpic}[width=0.44\columnwidth]{MainRegions.pdf} %,grid=true
	\put(-6,85){\textbf{b}}
	\end{overpic}
	~~
	\begin{overpic}[width=0.48\columnwidth]{TransferRegions.pdf}
	\put(-6,80){\textbf{c}}
	\end{overpic}
\end{center}
\vspace{-1.em}
\caption{\label{fig:Region}  Outlier rejection state machine and regions.
}
\vspace{-1.5em}
\end{figure}

@@

%\begin{figure}
%%  \vspace{-20pt}
%  \begin{center}
%\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\end{overpic}
%  \end{center}
%%  \vspace{-1em}
%\caption{\label{fig:Shapes} The six equal-area objects tested in simulation. %Six object shapes were tested.
%\vspace{-1em}
%}
%\end{figure}

@@

%\subsection{Automated object manipulation (simulation)}
 Fig.~\ref{fig:story} shows snapshots during an execution of this algorithm in simulation. Fig.~\ref{fig:DifferentMazes}a shows the six equal-area object shapes tested in simulations.
 To illustrate the flexibility of the algorithm we tested two additional workspaces, \emph{{\sffamily E}-shaped} and  \emph{Spiral}, without changing the algorithm. These are shown in Fig.~\ref{fig:Emaze}. 
 More complicated workspaces could be generated by composing these workspaces. 
 Fig.~\ref{fig:DifferentMazes}b shows the results of all  three mazes. The {\sffamily E}-shaped maze required the least average time because the path to the goal is shorter. 
   Experimental results of parameters sweeps are summarized in Fig.~\ref{fig:AutoVeryParam}.  Each trial measured the time to deliver the object to the goal location.  The default parameter settings used 100 particles, a normalized weight of 1, a hexagon shape, and Brownian noise (applied once each simulation step) with $W=5$.

@@

\begin{overpic}[width =\columnwidth]{SwarmRun.pdf}\end{overpic}
\vspace{-2em}
\caption{\label{fig:story}Snapshots showing an object manipulation simulation with 100 particles under automatic control (see also Extension 1).
\vspace{-2em}
}
\end{figure}

\begin{figure}\begin{center}\begin{overpic}[width=0.3\columnwidth]{MDPEmaze.pdf}\end{overpic}~\begin{overpic}[width=0.3\columnwidth]{EmazeFinal.pdf}\end{overpic}\\
\vspace{.25em}
\begin{overpic}[width=0.3\columnwidth]{Spiral.pdf}\end{overpic}~\begin{overpic}[width=0.3\columnwidth]{SpiralSuccess.pdf}\end{overpic}
\end{center}
\vspace{-1.em}
\caption{\label{fig:Emaze} We tested three workspaces. The control policies from value iteration for an {\sffamily E}-shaped and a spiral workspace are shown in the left column.
}
\end{figure}
\begin{figure}
\centering
\begin{overpic}[width=0.4\columnwidth]{AllShapes.pdf}\put(-6,85){\textbf{a}}\end{overpic}
\begin{overpic}[width=.5\columnwidth]{DifferentShapes.pdf}\put(0,68){\textbf{b}}\end{overpic}
\vspace{-0.5em}
\caption{\label{fig:DifferentMazes} a. The six equal-area objects tested in simulation. b. Completion times for three workspaces.}
\end{figure}

@@

\begin{figure*}
\centering
\renewcommand{\figwid}{0.5\columnwidth}
\begin{overpic}[width =\figwid]{SimVeryNum.pdf}\put(1,55){a)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryNoise.pdf}\put(1,55){b)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryWeight.pdf}\put(1,55){c)}
\end{overpic}
\begin{overpic}[width =\figwid]{SimVeryShape.pdf}\put(1,55){d)}
\end{overpic}
\vspace{-2em}
\caption{\label{fig:AutoVeryParam}Parameter sweep simulation studies for a) number of particles, b) different noise values, c) object weight, and d) object shape.  Each bar is labelled with the number of trials. Completion time is in seconds.
%\vspace{-2em}
}
\end{figure*}
